skating

### 2021-01-30 inst/skating.Rmd
consistentification

title: "Ladies' figure skating"
author: "Robin K. S. Hankin"
date: "`r Sys.Date()`"
vignette: >
  %\VignetteIndexEntry{figure_skating}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

@@

```{r, echo=T, results = 'hide', message= FALSE}
library("hyper2",quietly=TRUE)
```

This document analyses strengths of the competitors in the Ladies'
Figure Skating at the 2002 Winter Olympics.  In the package, dataframe
`skating_table` is copied from Lock and Lock (see skating.Rd for more
details).  It also creates file `data/skating.rda` which contains R objects `skating_table`,

```{r,showskating}
set.seed(0)
skating_table <- read.table("skating.txt",header=TRUE)
```

Object `skating_table` is an **order table**.  It is structured so
that each competitor is a row, and each judge is a column.  Function
`rank_likelihood()` considers each row to be a race [or a judge], so
we need to take a transpose.  The following R idiom is plausible but
incorrect:

```{r,skatingwrong}
skating_incorrect <- rank_likelihood(t(skating_table))
```

The above is incorrect because we are applying a rank technique to an
order table.  Correct analysis follows; firstly we coerce to a
**rank** table:

```{r,coercetoranktable}
(skating_order <- ordertable_to_ranktable(skating_table))
```

In the above, the column names should be read "came first" (`c1`),
"came second" (`c2`) and so on to "came 23rd" (`c23`).  From the first
column, we see that Hughes came first 5 times and Slutskya came first
four times, and from the last column we see that all judges except
`J3` awarded Luca last place.  File `man/rrank.Rd` has a detailed
discussion of the differences between rank and order.

@@

In the above, each row is a competitor and each column corresponds to
a ranking.  Thus the first column corresponds to "first place" and
shows that five judges ranked Hughes first and four judges ranked
Slutskaya first.  The second column corresponds to "second place" and
shows that one judge gave Hughes second place, two judges placed
Slutskaya second, five gave Kawn second, and one gave Cohen second.

@@

```{r,dpi=72}
 rL <- sort(skating_maxp,decreasing=TRUE)
 rL[] <- seq_along(rL)
 rO <- seq_len(nrow(skating_table))
 names(rO) <- rownames(skating_table)
 ordertransplot(rO,rL,xlab="offical rank",ylab="likelihood rank",main="Ladies free skating, 2002 Winter Olympics")
```

In the above diagram, there is close agreement between the two methods
in the large, but differences in detail.  For example, the official
ordering is hughes first, then slutskaya second, then kwan third;
likelihood winner is slutskya, followed by hughes and then kwan.

We can gain some understanding of this result by looking at the raw
judging results for the three medallists:

@@

```{r}
o <- maxp(o)
pie(o)
```

### Package dataset

>>>>

---
title: "skating"
output: bookdown::html_document2
---

```{r setup, include=FALSE}
set.seed(0)
knitr::opts_chunk$set(echo = TRUE)
library("hyper2")
library("magrittr")
options("digits" = 5)
```

@@

This document analyses strengths of the competitors in the Ladies'
Figure Skating at the 2002 Winter Olympics.  In the package, dataframe
`skating_table` is copied from Lock and Lock (see skating.Rd for more
details).  It also creates file `data/skating.rda` which contains R
objects `skating`, `skating_table`, and `skating_maxp`.

```{r,showskating}
set.seed(0)
skating_table <- read.table("skating.txt",header=TRUE)
skating_table
```

Object `skating_table` is an order table.  It is structured so that
each competitor is a row, and each judge is a column.  Function
`rank_likelihood()` considers each row to be a race [or a judge], so
we need to take a transpose.  The following R idiom is plausible but
**incorrect**:

```{r,skatingwrong}
skating_incorrect <- rank_likelihood(t(skating_table)) # incorrect
```

The above is incorrect because we are applying a rank technique to an
order table.  Correct analysis follows.  One good thing to do is to
present the dataset as an order table;

```{r,coercetoranktable}
ordertable_to_ranktable(skating_table)
```

In the above, the column names should be read "came first" (`c1`),
"came second" (`c2`) and so on to "came 23rd" (`c23`).  From the first
column, we see that Hughes came first 5 times [according to judges
1,5,7,8,9] and Slutskya came first four times [judges 2,3,4,6].  From
the last column we see that all judges except `J3` awarded Luca last
place.  File `man/rrank.Rd` has a detailed discussion of the
differences between rank and order.

@@

In the above, each row is a competitor and each column corresponds to
a ranking.  Thus the first column corresponds to "first place" and
shows that five judges ranked Hughes first and four judges ranked
Slutskaya first.  The second column corresponds to "second place" and
shows that one judge gave Hughes second place, two judges placed
Slutskaya second, five gave Kawn second, and one gave Cohen second.
The first row corresponds to Hughes and we see that, of $5+1+1+2=9$
judges, 5 placed Hughes first, one placed her second, one placed her
third, and two placed her fourth.

@@

```{r rLrO,fig.cap="Official ranks vs likelihood rank, Ladies free skating, 2002 Winter Olympics"}
 rL <- sort(skating_maxp,decreasing=TRUE)
 rL[] <- seq_along(rL)
 rO <- seq_len(nrow(skating_table))
 names(rO) <- rownames(skating_table)
 ordertransplot(rO,rL,xlab="offical rank",ylab="likelihood rank",main="Ladies free skating, 2002 Winter Olympics")
```

In figure \@ref(fig:rLrO), there is close agreement between the two
methods in the large, but differences in detail.  For example, the
official ordering is hughes first, then slutskaya second, then kwan
third; likelihood winner is slutskya, followed by hughes and then
kwan.  We can gain some understanding of this result by looking at the
raw judging results for the three medallists:

@@

```{r}
(o <- maxp(o))
pie(o)
```

### Package dataset {-}
